---
title: "owidR"
author: Piers York
output: github_document
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "images/"
)

library(dplyr)
```

This package acts as an interface to [Our World in Data](https://ourworldindata.org/) datasets, allowing for an easy way to search through data used in over 3,000 charts and load them into the R environment.

**Warning:** Package has recently been updated to remove `owid_get_datasets()` and change how the syntax works. It's now simpler to use but not backwards compatible.

## Installation

``` {r eval = FALSE}
devtools::install_github("piersyork/owidR")
```

## Using the package
The main function in owidR is `owid()`, which takes a chart id and returns a tibble (dataframe) of the corresponding OWID dataset. To search for chart ids you can use `owid_search()` to list all the chart ids that match a keyword or regular expression.

## Example
Lets use the core functions to get data on how human rights have changed over time. First by searching for charts on human rights.
  
``` {r example, eval = TRUE, message = FALSE}
library(owidR)

owid_search("human rights")

```

Let's use the human rights scores dataset.

``` {r example2, eval = TRUE, message = FALSE}

rights <- owid("human-rights-scores")

rights

```


`owid_plot()` makes it easy to visualise an owid dataset, plotting the first value column of an owid dataset. By default the mean score across all countries is plotted.
```{r owid_plot, eval = TRUE, results='hide'}
owid_plot(rights)
```

Use `summarise = FALSE` to show individual countries instead of the mean score. Unless a vector of entities is specified using the `filter` argument 9 random entities will be plotted. If the data is not a time-series then a bar chart of the entities values will be plotted.
```{r owid_plot2, eval = TRUE, results='hide'}
owid_plot(rights, summarise = FALSE, filter = c("North Korea", "South Korea", "France", "United Kingdom", "United States"))
```

`owid_map()` makes it easy to create a choropleth world map of datasets that contain country level data. The Entities of the owid data must be country names. Currently the function plots data for the most recent year.
```{r map, results='hide'}
owid_map(rights)
```

## Creating Our World in Data style graphs
`owid_grapher()` creates graphs in the style of Our World in Data. The output of `owid_grapher()` can be piped into `grapher_line()` to add a line graph, into `grapher_map()` to add a world map, and into `grapher_labels()` to add labels to the graph. The graph is shown in the RStudio viewer, or when called in an RMarkdown html document is displayed within the document.
```{r grapher_eg, results='hide'}
rights %>% 
  owid_grapher(x = year, y = `Human Rights Score (Schnakenberg & Fariss, 2014; Fariss, 2019)`, 
               entity = entity) %>% 
  grapher_line(selected = c("North Korea", "South Korea", "France", "United Kingdom", "United States")) %>% 
  grapher_map(palette = "RdYlGn", bins = c(-2, 0, 2, 4)) %>% 
  grapher_labels(title = "Human Rights Scores",
                 subtitle = "Values range from around -3.8 to around 5.4 (the higher the better)",
                 source = "Our World in Data; Schnakenberg and Fariss (2014); Fariss (2019)")
```





