---
title: "owidR"
author: Piers York
output: github_document
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "images/"
)
library(owidR)
```

This package acts as an interface to the [Our World in Data](https://ourworldindata.org/) datasets [GitHub repository](https://github.com/owid/owid-datasets), allowing for easy way to search from a list of almost 1,000 datasets and load them into the R environment.

## Installation

``` {r eval = FALSE}
devtools::install_github("piersyork/owidR")
```

## Using the package
There are three core functions in the `owidR` package. `owid_get_datasets()` returns a tibble of all available Our World in Data datasets alongside a generated id. `owid_search()` makes it easy to search through the available datasets. `owid()` takes an id and returns the corresponding dataset.

### Example
Get Covid-19 policy stringency data using the three core commands
``` {r example, eval = TRUE, message = FALSE}

ds <- owid_get_datasets()

id <- owid_search(ds, "COVID Government Response ")$id

covid_policy <- owid(ds, id)

covid_policy

```

Easily plot the first value column of the dataset using `owid_plot()`.
```{r owid_plot, eval = TRUE, results='hide'}
owid_plot(covid_policy)
```



