---
title: "owidR"
output: github_document
always_allow_html: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "inst/images/", fig.width=8.5, fig.asp=0.55, fig.align='center',
  message = FALSE
)

library(ggplot2)

ggplot2::theme_set(
  theme_minimal(base_family = "serif") %+replace%
    theme(text = element_text(colour = "#373737", size = 18),
          plot.title = element_text(size = "22", hjust = 0, vjust = 3.2, face = "bold"),
          plot.subtitle = element_text(hjust = 0, vjust = 3.5),
          plot.title.position = "plot",
          plot.caption.position = "plot",
          plot.caption = element_text(hjust = 0, size = 14, vjust = -1),
          legend.position = "right",
          axis.text = element_text(face = "plain"),
          axis.title.x = element_blank(),
          legend.title = element_blank(),
          panel.grid.major = element_line(linetype = "dashed"),
          panel.grid.minor = element_blank(),
          plot.margin = margin(11, 15, 8, 0),
          axis.line.x = element_line(colour = "#8e8e8e"),
          axis.ticks = element_line(colour = "#8e8e8e"))
)

```

<!-- badges: start -->
![CRAN downloads](http://cranlogs.r-pkg.org/badges/grand-total/owidR)
[![CRAN status](https://www.r-pkg.org/badges/version/owidR)](https://CRAN.R-project.org/package=owidR)
[![R-CMD-check](https://github.com/piersyork/owidR/workflows/R-CMD-check/badge.svg)](https://github.com/piersyork/owidR/actions)
<!-- badges: end -->

This package acts as an interface to [Our World in Data](https://ourworldindata.org/) datasets, allowing for an easy way to search through data used in over 3,000 charts and load them into the R environment.

## Installation
To install from CRAN:
```{r eval=FALSE}
install.packages("owidR")
```

To install the development version from GitHub: 
``` {r eval = FALSE}
devtools::install_github("piersyork/owidR")
```

## Using the package
The main function in owidR is `owid()`, which takes a chart id and returns a tibble (dataframe) of the corresponding OWID dataset. To search for chart ids you can use `owid_search()` to list all the chart ids that match a keyword or regular expression.

## Example
Lets use the core functions to get data on how human rights have changed over time. First by searching for charts on human rights.
  
``` {r example, eval = TRUE, message = FALSE}
library(owidR)

owid_search("human rights")

```

Let's use the human rights protection dataset.

``` {r example2, eval = TRUE, message = FALSE}

rights <- owid("human-rights-protection")

rights

```


ggplot2 makes it easy to visualise our data.
```{r plot}
library(ggplot2)
library(dplyr)

rights |> 
  filter(entity %in% c("United Kingdom", "France", "United States")) |> 
  ggplot(aes(year, `Human rights protection`, colour = entity)) +
  geom_line()
```


## COVID-19 Data
You can quickly download world covid-19 data, including vaccination rates, using `owid_covid()`. 

```{r covid}
covid <- owid_covid()

covid
```

## To-do
* [ ] Add function to load multiple country datasets into on dataframe
* [ ] Add caching of data (inc. backend)
* [ ] Remove interactive plotting to reduce dependencies
* [ ] Create way to import owid explorers



