---
title: "owidR"
author: Piers York
output: github_document
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "images/"
)
library(owidR)
```

This package acts as an interface to the [Our World in Data](https://ourworldindata.org/) datasets [GitHub repository](https://github.com/owid/owid-datasets), allowing for an easy way to search from a list of almost 1,000 datasets and load them into the R environment.

## Installation

``` {r eval = FALSE}
devtools::install_github("piersyork/owidR")
```

## Using the package
There are three core functions in the `owidR` package. `owid_get_datasets()` returns a tibble of all available Our World in Data datasets alongside a generated id. `owid_search()` makes it easy to search through the available datasets, using either a keyword or a regular expression. `owid()` takes an id and returns the corresponding dataset.

### Example
Lets use the core functions to get data on how human rights have changed over time.   
   
First loading the tibble of datasets.
  
``` {r example, eval = TRUE, message = FALSE}

ds <- owid_get_datasets()

```

Then searching for datesets about human rights.

``` {r example2, eval = TRUE, message = FALSE}

owid_search(ds, "human rights")
```

Let's use the Schnakenberg and Fariss (2014) dataset.

``` {r, example3, eval = TRUE, message = FALSE}

id <- owid_search(ds, "Human Rights Scores – Schnakenberg and Fariss")$id

rights <- owid(id, ds)

rights

```

`owid_plot()` makes it easy to visualise an owid dataset, plotting the first value column of an owid dataset. By default the mean score across all countries is plotted.
```{r owid_plot, eval = TRUE, results='hide'}
owid_plot(rights)
```

Use `summarise = FALSE` to show individual countries instead of the mean score. Unless a vector of entities is specified using the `filter` argument 9 random entities will be plotted. If the data is not a time-series then a bar chart of the entities values will be plotted.
```{r owid_plot2, eval = TRUE, results='hide'}
owid_plot(rights, summarise = FALSE, filter = c("North Korea", "South Korea", "France", "United Kingdom", "United States"))
```

`owid_map()` makes it easy to create a choropleth world map of datasets that contain country level data. The Entities of the owid data must be country names. Currently the function plots data for the most recent year.
```{r map, results='hide'}
owid_map(rights)
```

